Silenced in the Chamber: Analyzing Gendered Speech Interruptions in the Australian Parliament (1998–2022)

### 2025-03-28 paper/paper.qmd
add appendix, more fixes

```{r}
library(tinytex)

@@

xelatex --version

@@

library(tidyverse)
library(arrow)
library(here)
library(rstanarm)
library(modelsummary)
library(ggplot2)
library(knitr)
library(tibble)
library(gt)
library(patchwork)
library(kableExtra)
library(dplyr)
library(stringr)
library(stm)
library(quanteda)
library(tm)
```

@@

@preview_cleaned provides a snapshot of key variables, including year, gender, interject, party, and word_count, allowing us to inspect how the dataset is structured post-cleaning. @sum_stats presents aggregated counts of speeches by gender, interruption status, and party affiliation. It highlights the number of speeches given by male and female parliamentarians, how often interruptions occur, and the distribution of party representation. Missing values are explicitly recorded where relevant.
```{r}
#| echo: false
#| warning: false
#| message: false
#| tbl-cap: "Preview of the Cleaned Hansard Dataset"
#| label: tbl-preview_cleaned
# Select important variables for preview
preview_table <- hansard_corpus %>%
  select(year, gender, interject, party, word_count, fedchamb_flag) %>%
  head(6)  # Show only the first 6 rows

@@

\begin{align}
y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
\mbox{logit}(\pi_i) &= \alpha + \beta_1 \times \mbox{gender}_i + \beta_2 \times \mbox{party}_i + \beta_3 \times \mbox{word\_count}_i + \beta_4 \times \mbox{word\_count\_gender}_i \nonumber \\
\alpha &\sim \mbox{Normal}(0, 5) \nonumber \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \nonumber \\
\beta_2 &\sim \mbox{Normal}(0, 2.5) \nonumber \\
\beta_3 &\sim \mbox{Normal}(0, 2.5) \nonumber \\
\beta_4 &\sim \mbox{Normal}(0, 2.5) \nonumber
\end{align}

In this model, \( y_i \) represents the binary outcome variable indicating whether a speech was interrupted (\(\text{interject} = 1\)) or not (\(\text{interject} = 0\)). The probability of an interruption (\(\pi_i\)) is modeled using a logistic link function, where the log-odds of an interruption are expressed as a linear combination of the intercept (\(\alpha\)) and the predictor variables: where \(\beta_1, \beta_2, \beta_3, \beta_4\) are the regression coefficients corresponding to each predictor. The variables included in the model are defined as follows: \(\text{gender}_i\) represents the speaker’s gender (Male, Female, or Unknown), \(\text{party}_i\) indicates the political party affiliation, \(\text{word\_count}_i\) captures the total number of words spoken in a given speech, and \(\text{word\_count\_gender}_i\) represents the interaction between word count and gender.

To regularize the model and incorporate prior information, we assign informative normal priors to the intercept and coefficients. Specifically, the intercept \(\alpha\) follows a normal distribution with a mean of 0 and a standard deviation of 5, while the coefficients \(\beta_1, \beta_2, \beta_3,\) and \(\beta_4\) follow normal distributions with a mean of 0 and a standard deviation of 2.5. These priors help stabilize the estimates, particularly in cases where the data might be sparse or imbalanced.

@@

@@

@tbl-modelresults1 identifies key predictors including gender, party, and word count of speech interruptions in parliamentary proceedings. Gender plays an important role, with male speakers more likely to experience interruptions compared to female speakers. This suggests that men face slightly higher odds of interjections during their speeches. Party affiliation also significantly influences interruptions, with some parties, such as the ALP, experiencing more frequent interjections, while others, like the LNP, are interrupted less often. These patterns highlight how political dynamics can shape interactions in debates.

@@

Model diagnostics, including convergence checks and posterior summaries, and explanations can be found in @Appendix

@@

### Model justification

We expect a positive relationship between the= size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

@@

```{r}
library(bayesplot)
#mcmc_trace(as.array(model1))

```{r}
#pp_check(model1, nsamples = 100)

# Additional data details

@@

## Diagnostics

>>>>

@@

@@

library(tidyverse)
library(arrow)
library(here)
library(rstanarm)
library(modelsummary)
library(ggplot2)
library(knitr)
library(tibble)
library(gt)
library(patchwork)
library(kableExtra)
library(dplyr)
library(stringr)
library(stm)
library(quanteda)
library(tm)
library(tinytex)
library(bayesplot)

@@

@tbl-preview_cleaned provides a snapshot of key variables, including year, gender, interject, party, and word_count, allowing us to inspect how the dataset is structured post-cleaning. @tbl-sum_stats presents aggregated counts of speeches by gender, interruption status, and party affiliation. It highlights the number of speeches given by male and female parliamentarians, how often interruptions occur, and the distribution of party representation. Missing values are explicitly recorded where relevant.
```{r}
#| echo: false
#| warning: false
#| message: false
#| tbl-cap: "Preview of the Cleaned Hansard Dataset"
#| label: tbl-preview_cleaned
# Select important variables for preview
preview_table <- hansard_corpus %>%
  select(year, gender, interject, party, word_count, fedchamb_flag) %>%
  head(6)  # Show only the first 6 rows

@@

\begin{align}
y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
\mbox{logit}(\pi_i) &= \alpha + \beta_1 \times \mbox{gender}_i + \beta_2 \times \mbox{party}_i + \beta_3 \times \mbox{word\_count}_i + \beta_4 \times \mbox{word\_count\_gender}_i \nonumber \\
\alpha &\sim \mbox{Normal}(0, 5) \nonumber \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \nonumber \\
\beta_2 &\sim \mbox{Normal}(0, 2.5) \nonumber \\
\beta_3 &\sim \mbox{Normal}(0, 2.5) \nonumber \\
\beta_4 &\sim \mbox{Normal}(0, 2.5) \nonumber
\end{align}
In this model, $y_i$ represents the binary outcome variable indicating whether a speech was interrupted ($\text{interject} = 1$) or not ($\text{interject} = 0$). The probability of an interruption ($\pi_i$) follows a logistic distribution:

$\text{logit}(\pi_i) = \alpha + \beta_1 \times \text{gender}_i + \beta_2 \times \text{party}_i + \beta_3 \times \text{word\_count}_i + \beta_4 \times \text{word\_count\_gender}_i$

where $\beta_1, \beta_2, \beta_3,$ and $\beta_4$ are the regression coefficients. The model includes the following predictors: $\text{gender}_i$ (Male, Female, or Unknown), $\text{party}_i$ (political party affiliation), $\text{word\_count}_i$ (total words spoken in a speech), and $\text{word\_count\_gender}_i$ (interaction between word count and gender).

To regularize the model, we assign normal priors to the intercept and coefficients. The intercept $\alpha$ follows a normal distribution with mean 0 and standard deviation 5, while the coefficients $\beta_1, \beta_2, \beta_3,$ and $\beta_4$ follow normal distributions with mean 0 and standard deviation 2.5. These priors provide stability, especially when data is sparse or imbalanced.

@@

## Model justification

A relationship is expected between gender, party affiliation, and speech interruptions in parliamentary debates. Male speakers are likely to face more interruptions than female speakers, reflecting common patterns in political discourse. However, interruptions may still occur for female speakers depending on the context of their speech.

For party affiliation, the ALP and LNP, as the two main parties, are expected to show different interruption patterns. ALP members may experience more interruptions, especially when in opposition, while LNP members may face fewer interruptions, particularly when in government. Other parties and independents may not follow a clear pattern due to their varying roles in debates.

Speech length also matters—longer speeches may be interrupted less, possibly because they allow speakers to establish their points, while shorter speeches might invite more interjections. The interaction between word count and gender suggests that longer speeches by male speakers could attract more interruptions, while this trend may be weaker for female speakers.

A Bayesian approach helps capture uncertainty in these patterns while ensuring stable estimates, even with imbalanced data. This method provides a clearer view of how gender, party, and speech characteristics influence interruptions in parliamentary debates.

@@

@tbl-modelresults1 identifies key predictors including gender, party, and word count of speech interruptions in parliamentary proceedings. The intercept, which represents the baseline category in the model, corresponds to female speakers from the LP party. This suggests that all other coefficients should be interpreted as deviations from this baseline.Gender plays an important role, with male speakers more likely to experience interruptions compared to female speakers. This suggests that men face slightly higher odds of interjections during their speeches. Party affiliation also significantly influences interruptions, with some parties, such as the ALP, experiencing more frequent interjections, while others, like the LNP, are interrupted less often. These patterns highlight how political dynamics can shape interactions in debates.

@@

Model diagnostics, including convergence checks and posterior summaries, and explanations can be found in Appendix @sec-model-details

@@

@@

@@

```{r}
#| label: fig-post_dist
#| fig-cap: Posterior distribution
#| echo: false
#| warning: false
#| message: false
pp_check(model1, nsamples = 100) +
  theme(legend.position = "bottom")

```
In @fig-post_dist, a posterior distribution is implemented. This compares the actual instances of speech interruptions with the predicted probabilities from the posterior distribution of the Bayesian logistic regression model. The close alignment between the posterior distribution and the observed data suggests that my model provides a strong representation of the factors influencing interruptions in parliamentary speech.

## Markov chain Monte Carlo Convergence Check

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-trace
#| fig-cap: "Trace plot"
#| fig.width: 12
#| fig.height: 25   # Increased height for better spacing

# Generate trace plot with better spacing
trace_plot <- mcmc_trace(
  as.array(model1),
  facet_args = list(ncol = 2)  # Keep two plots per row
) +
  theme_minimal(base_size = 14) +  # Increase base text size
  theme(
    axis.text.x = element_text(size = 12),  
    axis.text.y = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),  
    legend.text = element_text(size = 12),  
    legend.title = element_text(size = 14),
    panel.spacing.y = unit(2, "cm")  # **Increase vertical spacing**
  )

trace_plot

@fig-trace is the trace plot of the model. The lines are horizontal and oscillating, and have overlaps between the chains, suggesting no existing signs of issues with the model.

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-rhat
#| fig-cap: "Rhat plot"
plot(model1, "rhat")

```
@fig-rhat is the Rhat plot of the model. It compares the variability within each chain to the variability between chains in MCMC. In ideal condition, the Rhat should be close to 1, indicating that the chains have converged to the same distribution. This Rhat plot is indeed close to 1, suggesting that the MCMC algorithm has reached convergence for the model. 
## Credibility Interval {#sec-credibility-interval}

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-modelresults1
#| fig-cap: "Credible intervals for predictors of interruptions"

modelplot(model1, conf_level = 0.90, size = 0.2) +
  labs(x = "90% credibility interval")

```
@fig-modelresults1 displays the 90% credibility intervals for the posterior distributions of the logistic regression model predicting interruptions in parliamentary speech. Each point represents the posterior mean, while the horizontal lines indicate the uncertainty in the estimates.

Although many party coefficients include zero within their credibility intervals—suggesting they do not have a statistically significant effect—the ALP and LNP show meaningful effects. These two parties, the dominant political groups in Australia, have coefficients that suggest they significantly influence interruptions. LP is included in the intercept, meaning all other party estimates should be interpreted relative to this baseline category.

Gender also plays a role in predicting interruptions. Both "Male" and "Female" categories show significant effects, while "Gender Unknown" does not, likely due to a smaller sample size. The strong effect for male speakers suggests that gender influences interjections, aligning with findings that men experience more interruptions in parliamentary debates.

Additionally, word count has a noticeable effect. Longer speeches tend to be associated with fewer interruptions, implying that shorter speeches might invite more interjections. The interaction term between word count and gender suggests that male speakers might experience a slight increase in interruptions as speech length grows, whereas the effect is less pronounced for female speakers.
